HyperLogLog: Exponentially Bad in Adversarial Settings

### 2022-02-14 docs/HLL_exp_bad/main.tex
redis graph

\begin{figure*}[tb!]
\centering
     \makebox[\linewidth]{
\fbox{
\begin{minipage}{40em}
\begin{flushleft}
\textit{Let $h : \hat{\mathcal{D}} \rightarrow \{0, 1\}^{32}$ hash data from some domain $\hat{\mathcal{D}}$ to binary $32$-bit strings.\\
For $s$ of bit-length $\ell$, let $\rho(s)$ denote the position of the leftmost $1$-bit of $s$: e.g., $\rho(1...) = 1, \rho(0001...) = 4, \rho(0^\ell) = \ell + 1.$\\}
\textbf{define} $m = 2^n$ with $n \in [4..16].$\\
\textbf{define} $\alpha_{16} = 0.673$; $\alpha_{32} = 0.697$; $\alpha_{64} = 0.709$; $\alpha_{m} = 0.7213/(1+ 1.079/m)$ for $m \ge 128$;\\
\vspace{0.5em}
\textbf{Algorithm} HyperLogLog (\textbf{input} ${\mathcal{X}}$ : multiset of items from domain $\hat{\mathcal{D}}$).\\
\textbf{initialize} a collection of $m$ registers, $M[1], ..., M[m]$, to $0$;\\
\vspace{0.5em}
\textbf{for} $x \in {\mathcal{X}}$ \textbf{do}\\
    \hspace{2em} \textbf{set} $y := h(x);$\\
    \hspace{2em} \textbf{set} $j = 1 + <y_1y_2...y_{n}>_2$; \hfill \textit{\{bucket index determined by the first $n$ bits of $y$\}}\\
    \hspace{2em} \textbf{set} $w := y_{n+1}y_{n+2} \ldots y_{n+\ell}$;\hfill \textit{\{bit string determined by the last $\ell$ bits of $y$\}}\\
   \hspace{2em} \textbf{set} $M[j] :=$ max$(M[j], \rho(w))$;\\
\vspace{0.2em}
\textbf{compute} $E := \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$; \hfill \textit{\{the raw HyperLogLog estimate\}}\\
\vspace{0.2em}
\textbf{if} $E \le \frac{5}{2}m$ \textbf{then}\\
    \hspace{2em}  \textbf{let} $V$ be the number of registers equal to $0$;\\
    \hspace{2em}  \textbf{if} $V \neq 0$ \textbf{then set} $E^*:=  m\cdot \ln(m/V)$ \textbf{else set} $E^*:= E$; \hfill \textit{\{small range correction\}}\\
\textbf{if} $E \le \frac{1}{30}2^{32}$ \textbf{then}\\
    \hspace{2em} \textbf{set} $E^*:= E$; \hfill \textit{\{intermediate range --- no correction\}}\\
\textbf{if} $E > \frac{1}{30}2^{32}$ \textbf{then}\\
    \hspace{2em}  \textbf{set} $E^*:= -2^{32}\ln(1 - E/2^{32})$; \hfill \textit{\{large range correction\}}\\
\textbf{return} \textit{cardinality estimate $E^*$ with typical relative error} $\pm1.04/\sqrt{m}$.
\end{flushleft}
\end{minipage}
}}
\caption{The HyperLogLog Algorithm from~\cite{hll}}
\label{fig:hll}
\end{figure*}

@@

We assume that at least $2^{t-1} m\ln m$ items have already been honestly inserted into the target HLL, for some integer $t$ with $1 \leq t \leq \ell$ which will be a parameter of our attack. We assume the S2 adversary somehow knows that this condition holds (e.g.\ through knowledge of when the HLL was initialised and the average rate at which items are inserted into it). Notice that an S3 adversary can use its ability to get cardinality estimates to obtain this knowledge; this is one situation where an S3 adversary can make good use of the extra information available to it as compared to an S2 adversary. In S4, the adversary can use its access to a snapshot to mount a more powerful attack; see Section~\ref{sec:nonemptyHLLS4}.

@@

\begin{table*}[tb!]
\caption{Comparison of attack results in the S1 adversarial scenario, in the setting of~\cite{hllvuln}.}
\begin{center}
\begin{tabular}{| l | >{\raggedleft\arraybackslash}p{6.8em} | >{\raggedleft\arraybackslash}p{6.8em} | >{\raggedleft\arraybackslash}p{6.8em} |}
    \hline
    \textbf{Attack} & \multicolumn{2}{c|}{Ours} &\multicolumn{1}{c|}{~\cite{hllvuln}} \\ \hline
      \textbf{Original Card. Est.} & 0 & 0 & 0  \\ \hline
       \textbf{\# Targeted buckets $B$} & 1 & 128 ($m/2$) & - \\ \hline
     \textbf{\# Items added} & 977 & 125,091 & 74,390 \\ \hline
    \textbf{Final Card. Est.} & 1 & 177 &  15,780 \\ \hline
    \textbf{\# Resets} & 1,148 & 1,655 & 3 \\ \hline

@@

\begin{figure*}[htp]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \caption{Attack results as function of t in scenario S2 with initially non-empty HLL sketch.}\label{fig:graphS2}
        \includegraphics[width=0.9\textwidth]{figures/graph_S2.pdf}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \caption{Attack results as function of t in scenario S4 with initially non-empty HLL sketch.}\label{fig:graphS4}
        \includegraphics[width=0.9\textwidth]{figures/graph_S4.pdf} % second figure itself
     \end{minipage}
\end{figure*}

@@

\subsection{Results for Redis}
We also implemented our attacks for the HLL implementation used in Redis as described in Section~\ref{sec:redis}.
The hash function $h$ is fixed as the 64-bit MurmurHash3 function. The seed used internally is hardcoded and can be extracted by inspection of the openly-available Redis codebase.
Furthermore, the number of buckets is constant and set to 16,384. Knowledge of these parameters places us in the setting of an S2 adversary.
Since a classic Redis application separates the server holding the sketch from the client interacting with the sketch, we consider that attack scenarios S3 or S4, while possible, are not typical for Redis. 
Thus, we only consider scenario S2 for Redis.
For simplicity, we set attack parameter $\epsilon$ to $\frac{1}{2}$ in the attack on an initially empty HLL sketch.
The results are averaged over 50 iterations and reported in Table~\ref{table:tab3}.
Similarly to the case for classic HLL, our empirical results match closely with our theoretical analysis.

\begin{table*}[tb!]
\begin{center}
\caption{Attack results for HLL in Redis, scenario S2, averaged over 50 iterations.}\label{table:tab3}
%\begin{tabular}{| m{11em} | m{4em} | m{4.2em} |}
\begin{tabular}{| c | r | r |}
    \hline
    \textbf{Original Card. Est.} & 0 & 139,967 \\ \hline
     \textbf{$\epsilon$} &  0.5 & - \\ \hline
    \textbf{\# Items added} & 524,210 & 524,454\\ \hline
    \textbf{Final Card. Est.} & 13,185 & 139,967 \\ \hline
\end{tabular}
\end{center}
\end{table*}
%\todo[inline]{Results from Redis or GBQ if we get them.}

>>>>

\begin{figure*}[tb!]
\centering
\caption{The HyperLogLog Algorithm from~\cite{hll}}
     \makebox[\linewidth]{
\fbox{
\begin{minipage}{40em}
\begin{flushleft}
\textit{Let $h : \hat{\mathcal{D}} \rightarrow \{0, 1\}^{32}$ hash data from some domain $\hat{\mathcal{D}}$ to binary $32$-bit strings.\\
For $s$ of bit-length $\ell$, let $\rho(s)$ denote the position of the leftmost $1$-bit of $s$: e.g., $\rho(1...) = 1, \rho(0001...) = 4, \rho(0^\ell) = \ell + 1.$\\}
\textbf{define} $m = 2^n$ with $n \in [4..16].$\\
\textbf{define} $\alpha_{16} = 0.673$; $\alpha_{32} = 0.697$; $\alpha_{64} = 0.709$; $\alpha_{m} = 0.7213/(1+ 1.079/m)$ for $m \ge 128$;\\
\vspace{0.5em}
\textbf{Algorithm} HyperLogLog (\textbf{input} ${\mathcal{X}}$ : multiset of items from domain $\hat{\mathcal{D}}$).\\
\textbf{initialize} a collection of $m$ registers, $M[1], ..., M[m]$, to $0$;\\
\vspace{0.5em}
\textbf{for} $x \in {\mathcal{X}}$ \textbf{do}\\
    \hspace{2em} \textbf{set} $y := h(x);$\\
    \hspace{2em} \textbf{set} $j = 1 + <y_1y_2...y_{n}>_2$; \hfill \textit{\{bucket index determined by the first $n$ bits of $y$\}}\\
    \hspace{2em} \textbf{set} $w := y_{n+1}y_{n+2} \ldots y_{n+\ell}$;\hfill \textit{\{bit string determined by the last $\ell$ bits of $y$\}}\\
   \hspace{2em} \textbf{set} $M[j] :=$ max$(M[j], \rho(w))$;\\
\vspace{0.2em}
\textbf{compute} $E := \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$; \hfill \textit{\{the raw HyperLogLog estimate\}}\\
\vspace{0.2em}
\textbf{if} $E \le \frac{5}{2}m$ \textbf{then}\\
    \hspace{2em}  \textbf{let} $V$ be the number of registers equal to $0$;\\
    \hspace{2em}  \textbf{if} $V \neq 0$ \textbf{then set} $E^*:=  m\cdot \ln(m/V)$ \textbf{else set} $E^*:= E$; \hfill \textit{\{small range correction\}}\\
\textbf{if} $E \le \frac{1}{30}2^{32}$ \textbf{then}\\
    \hspace{2em} \textbf{set} $E^*:= E$; \hfill \textit{\{intermediate range --- no correction\}}\\
\textbf{if} $E > \frac{1}{30}2^{32}$ \textbf{then}\\
    \hspace{2em}  \textbf{set} $E^*:= -2^{32}\ln(1 - E/2^{32})$; \hfill \textit{\{large range correction\}}\\
\textbf{return} \textit{cardinality estimate $E^*$ with typical relative error} $\pm1.04/\sqrt{m}$.
\end{flushleft}
\end{minipage}
}}

\label{fig:hll}
\end{figure*}

@@

We assume that at least $2^{t-1} m\ln m$ items have already been honestly inserted into the target HLL, for some integer $t$ with $1 \leq t \leq \ell$ which will be a parameter of our attack. We assume the S2 adversary somehow knows that this condition holds (e.g.\ through knowledge of when the HLL was initialised and the average rate at which items are inserted into it). If it were to be the case that our assumption does not hold and the adversary cannot estimate the number of items in the sketch, they can use a downward estimate such as $t=1$. As this lower-bound on $t$ is highly likely to happen, the adversary is taking only very small risk. Notice that an S3 adversary can use its ability to get cardinality estimates to obtain this knowledge; this is one situation where an S3 adversary can make good use of the extra information available to it as compared to an S2 adversary. In S4, the adversary can use its access to a snapshot to mount a more powerful attack; see Section~\ref{sec:nonemptyHLLS4}.

@@

\begin{table*}[tb!]
\caption{Comparison of attack results for a classic HLL implementation in the S1 adversarial scenario, \\ in the setting of~\cite{hllvuln}.}
\begin{center}
\begin{tabular}{| l | >{\raggedleft\arraybackslash}p{6.8em} | >{\raggedleft\arraybackslash}p{6.8em} | >{\raggedleft\arraybackslash}p{6.8em} |}
    \hline
    \textbf{Attack} & \multicolumn{2}{c|}{Ours} &\multicolumn{1}{c|}{~\cite{hllvuln}} \\ \hline
      \textbf{Original Card. Est.} & 0 & 0 & 0  \\ \hline
       \textbf{\# Targeted buckets $B$} & 1 & 128 ($m/2$) & - \\ \hline
     \textbf{\# Items added} & 977 & 125,091 & 74,390 \\ \hline
    \textbf{Final Card. Est.} & 1 & 177 &  15,780 \\ \hline
    \textbf{\# Resets} & 1,148 & 1,655 & 3 \\ \hline

@@

\begin{figure*}[htp]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \caption{Attack results as function of t in scenario S2 with initially non-empty HLL sketch.}\label{fig:graphS2}
        \includegraphics[width=0.9\textwidth]{figures/graph_S2.pdf}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \caption{Attack results as function of t in scenario S4 with initially non-empty HLL sketch.}\label{fig:graphS4}
        \includegraphics[width=0.9\textwidth]{figures/graph_S4.pdf} % second figure itself
     \end{minipage}

     \begin{minipage}{0.45\textwidth}
        \centering
        \caption{Attack results as function of t for HLL in Redis (scenario S2) with initially non-empty HLL sketch.}\label{fig:graphRedis}
        \includegraphics[width=0.9\textwidth]{figures/graph_redis.pdf} % second figure itself
     \end{minipage}
\end{figure*}

@@

\subsection{Results for Redis}
We also implemented our attacks for the HLL implementation used in Redis as described in Section~\ref{sec:redis}.
The hash function $h$ is fixed as the 64-bit MurmurHash3 function. The seed used internally is hardcoded and can be extracted by inspection of the openly-available Redis codebase.
Furthermore, the number of buckets is constant and set to 16,384. Knowledge of these parameters places us in the setting of an S2 adversary.
Since a classic Redis application separates the server holding the sketch from the client interacting with the sketch, we consider that attack scenarios S3 or S4, while possible, are not typical for Redis. 
Thus, we only consider scenario S2 for Redis.
We set attack parameter $\epsilon$ to $\frac{1}{2}$ and then $\frac{1}{4}$ in the attack on an initially empty HLL sketch. We recall that this will lead to the number of empty buckets $C_0$ to be equal to $\frac{m}{2} (=8,192)$ and $\frac{m}{4} (=4,096)$ respectively.
The results are averaged over 50 iterations and reported in Table~\ref{table:tab3} and Figure~\ref{fig:graphRedis}.
Similarly to the case for classic HLL, our empirical results match closely with our theoretical analysis.

\begin{table*}[tb!]
\begin{center}
\caption{Attack results on initially empty sketches for HLL in Redis (scenario S2), \\ averaged over 50 iterations.}\label{table:tab3}
%\begin{tabular}{| m{11em} | m{4em} | m{4.2em} |}
\begin{tabular}{| l | r | r |}
    \hline
    \textbf{Original Card. Est.} & 0 & 0 \\ \hline
     \textbf{$\epsilon$} &  0.5 & 0.25  \\ \hline
    \textbf{\# Items added} &  524,2104 & 785,920 \\ \hline
    \textbf{Final Card. Est.} & 13,185 & 35,938\\ \hline
\end{tabular}
\end{center}
\end{table*}
%\todo[inline]{Results from Redis or GBQ if we get them.}

%    \textbf{Original Card. Est.} & 0 & 139,967 \\ \hline
%     \textbf{$\epsilon$} &  0.5 & - \\ \hline
%    \textbf{\# Items added} & 524,210 & 524,454\\ \hline
%    \textbf{Final Card. Est.} & 13,185 & 139,967 \\ \hline
